---
title: "Class 15: Genotype Data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read sample genotypes data from ENSEMBL

We downloaded genotype data from ENSEMBL on the MXL Mexican Ancestry in Los Angeles, California dataset.

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

Q. What proportion of this population are G/G, etc?

We want to look at the second column that contains the genotype information

```{r}
table( mxl$Genotype..forward.strand. )
```

## RNA-seq result analysis for different genotypes of this SNP.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
# expr$geno == "G/G"
summary( expr[expr$geno == "G/G", ]$exp, breaks=20)
```

We'll use the `boxplot()` function and the input data will be **expr**. How do we draw a useful plot?

```{r}
boxplot(exp ~ geno, data=expr, notch=TRUE)
```

Q. How many individuals are in this data set?

```{r}
# alternatively, can view in global eivironment, # of observations / rows
nrow(expr)
```

