---
title: "Principal Component Analysis"
author: "Tiffany Chu"
date: "February 5, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The main function in base R for PCA is called `prcomp`. Here we will use PCA to examine the food that is eaten in the UK and in Northern Ireland.

Import the CSV file first (saved in project folder):
change the row heading names from the default (x) to the listed headings
```{r}
x <- read.csv("UK_foods.csv", row.names = 1)
x
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```


Show similarities in each region:
England on y-axis
duplicated information on bottom left
if they are the same, points will be on the line
  higher on the y-axis means English people more of that category compared to x-axis country
```{r}
pairs(x, col=rainbow(10), pch=16)
```

transpose (x) because base R takes the rows, so transposing will put countries in rows and categories as columns
```{r}
pca <- prcomp( t(x))
```

```{r}
summary(pca)
```

Summary (cumulative proportion) shows that 67% of data was captured in PC1, and 96.5% of the data was captured in PC1 and PC2.

What is in the PCA object?
```{r}
attributes(pca)
```
Printing just (pca) is not very useful:
```{r}
pca
```

However, we know from `attributes` that "x" is where the data we are interested in is, so we use pca$x

```{r}
plot(pca$x[,1], pca$x[,2])
```

Let's put in labels and color code the countries:
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], labels = colnames(x),
  col=c("black", "red", "blue", "darkgreen"))
```

To just get the score plot:
```{r}
plot(pca)
```

We can also project in new data (next class)