---
title: "Structural Bioinformatics Part I"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

> Q: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy?

```{r}
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats
```

```{r}
PDB_total <- sum(stats$Total)
PDB_total
```


```{r}
ans <- round(stats$Total / sum(stats$Total) * 100, 2)
names(ans) <- rownames(stats)
ans
```

> Q: What proportion of structures in the PDB are protein?

```{r}
protein <- round(sum(stats$Proteins) / sum(stats$Total) *100, 2)
protein
```


## Section 2. Using bio3d to analyze protein structure data

Use the bio3d package from CRAN.

```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb
```

## Section 3. Principal component analysis (PCA) on adenylate kinase

Perform PCA on the collection of adenylate kinase structures in the protein databank.


Adenylate kinase structures from the PDB for analysis

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A','4PZL_A')
```

Download and parse these structures

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

```{r}
# Align releated PDBs: fit = TRUE fits them on top of each other
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

```{r}
library(bio3d.view)
library(rgl)

view(pdbs)
```


```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```

```{r}
library(XML)
library(RCurl)
anno <- pdb.annotate(ids)
unique(anno$source)
anno
```

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

Make a PCA trajectory

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")

view(pc1)
```







