---
title: "Class 14: RNA-seq"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```
Q. How many genes are we dealing with?
```{r}
nrow(counts)
```

We will compare the control versus the treated count data contained in `counts`. We will first extract the control columns by looking up their colnames in the `metadata` object.

```{r}
# Accesss the "control" metadata IDs only

#metadata
#metadata$dex == "control"

control <- metadata[metadata$dex =="control", ]
control
```

```{r}
# Access the count columns with control$id
counts[ , control$id]
```

```{r}
# Find the mean value for each gene (i.e. each row)
control.mean <- rowSums(counts[ , control$id]) / nrow(control)
control.mean
```


```{r}
# Access the mean values of treated samples
treated <- metadata[metadata$dex == "treated", ]
counts[ , treated$id]
treated.mean <- rowSums(counts[ , treated$id]) / nrow(treated)
```

Control means versus treated means:

```{r}
mycounts <- data.frame(control.mean, treated.mean)
```

Plot data for quick check:
```{r}
plot(mycounts)
```

Heavily skewed data, so use log

```{r}
plot(mycounts, log="x")
```

It appears that we have some zero count genes
```{r}
View(mycounts)
```

Since we cannot determine anything from zero count genes in this dataset, we will remove them.

```{r}
to.rm <- 	unique(which(mycounts == 0, arr.ind = TRUE)[ , "row"])
newcounts <- mycounts[-to.rm, ]
```

Q. How many genes do we have now?

```{r}
nrow(newcounts)
```

Here we calculate log2foldchange, add it to our meancounts data.frame and inspect the results either with the `head()` or the `View()` function.

```{r}
newcounts$log2fc <- log2(newcounts[,"treated.mean"]/newcounts[,"control.mean"])
head(newcounts)
```

A common rule of thumb in the field is to use a log2fc of greater than +2 as *up-regulated* and less than -2 as *down-regulated*.

Q. How many of the genes in this dataset are up-regulated upon this drug treatment?

```{r}
sum(newcounts$log2fc > 2)
```

Q. How many genes are down-regulated at this threshold?

```{r}
sum(newcounts$log2fc < -2)
```

## Using DESeq2 for Analysis

```{r}
library(DESeq2)
```

Set up the object required by DEseq

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

Run the main DESeq calculation

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

## Volanco plots

Plot of log2fc vs p-value

```{r}
plot(res$log2FoldChange, -log(res$padj), col="gray")
abline(v=c(-2, +2), lty=2)
# abline(h=log(0.05), lty=2) new package has alpha as the p-value argument, put in alpha=0 to see all the data because otherwise alpha automatically applies p-value cutoff

```

