---
title: "Coronavirus Data Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coronavirus

Here we analyze infection data for the 2019 novel Coronavirus COVID-19 (2019-nCoV) epidemic. The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus repository.

A CSV file is available here https://github.com/RamiKrispin/coronavirus-csv


```{r}
url <- "https://tinyurl.com/COVID-2019"
virus <- read.csv(url)

tail(virus)
```

> Q1. How many total infected cases are there around the world?

```{r}
total_cases <- sum(virus$cases)
total_cases
```

Let's review the contents of the *$case* column

```{r}
table(virus$type)
```

> Q2. How many deaths linked to infected cases have there been?

```{r}
inds <- virus$type == "death"
death_cases <- sum(virus[inds, "cases"])
death_cases
```

> Q3. What is the overall death rate globally?

Percent  death

```{r}
round(death_cases/total_cases * 100, 2)
```

> Q4. What is the death rate in Mainland China?

```{r}
# Find all the entries for Mainland China
China.cases <- virus[virus$Country.Region == "Mainland China", ]

# Sum the total number of cases for Mainland China
sum.China.cases <- sum(China.cases$cases)

# Logical for deaths linked to infected cases for Mainland China
China.deaths <- China.cases$type == "death"

# Sum the total number of deaths linked to infected cases for Mainland China
sum.China.deaths <- sum(China.cases[China.deaths, "cases"])

# Percent death
round(sum.China.deaths/sum.China.cases * 100, 2)

```

> Q5. What is the death rate in Italy, Iran, and the United States?

For Italy

```{r}
Italy.cases <- virus[virus$Country.Region == "Italy", ]

Italy.deaths <- Italy.cases$type == "death"

# Percent death
round( (sum(Italy.cases[Italy.deaths, "cases"]) / sum(Italy.cases$cases)) * 100, 2)

```

For Iran

```{r}
Iran.cases <- virus[virus$Country.Region == "Iran", ]

Iran.deaths <- Iran.cases$type == "death"

# Percent death
round( (sum(Iran.cases[Iran.deaths, "cases"]) / sum(Iran.cases$cases)) * 100, 2)

```

For United States

```{r}
US.cases <- virus[virus$Country.Region == "US", ]

US.deaths <- US.cases$type == "death"

# Percent death
round( (sum(US.cases[US.deaths, "cases"]) / sum(US.cases$cases)) * 100, 2)
```

Function for death rate for country "x"

```{r}

coronavirus.death.rate <- function(x) {
  country.cases <- virus[virus$Country.Region == x, ]
  country.deaths <- country.cases$type == "death"
  round( (sum(country.cases[country.deaths, "cases"]) / sum(country.cases$cases)) * 100, 2)
}

# Example for function `coronavirus.death.rate()` for the death rate in the US
coronavirus.death.rate("US")
```

