---
title: "Structure Based Drug Design"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and begin processing structure

Here the HIV-Pr structure (PDB code: 1HSG) is downloaded from the main PDB database and cleaned up. Separate "protein-only" and "ligand only" PDB files will be created.

```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```

We use `read.pdb()`, `atom.select()` and `write.pdb()` functions to make the separate "protein-only" and "ligand only" PDB files.

```{r}
hiv <- read.pdb(file.name)
hiv
```

The protein-only file:
```{r}
prot <- atom.select(hiv, "protein", value = TRUE)
write.pdb(prot, file="1hsg_protein.pdb")
```

The ligand only file:
```{r}
lig <- atom.select(hiv, "ligand", value = TRUE)
write.pdb(lig, file="1hsg_ligand.pdb")
lig
```

## Read docking results

Read in the output of docking and make a PDB fie for viewing in VWD or PyMol
```{r}
res <-read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

