---
title: 'Class 06: R functions'
author: "Tiffany"
date: "January 24, 2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
## This is a level 2 heading.

This is text **bold** with more *text*

and a list of stuff with return carot

- fries
- truffles


```{r}
plot(1:5, col="blue", typ="o")
```

Let's insert a code chunk with the shortcut `Ctrl+Alt+i`:
```{r}
x <-  c(1:10)
x
```

Here is my analysis of your new data.  It looks ok...the mean of your data is `r mean(x)`.

## More on reading input files

We will use the read.table() function
```{r}
read.table("test1.txt")
```

We will use the read.table() function.
Look at arguments needed for this data:
```{r}
read.table("test1.txt", header = TRUE, sep=",")
x <- read.table("test1.txt", header = TRUE, sep=",")
```

Alternatively, use read.csv
```{r}
read.csv("test1.txt")
```

```{r}
x2 <-  read.table("test2.txt", header = TRUE, sep ="$")
x2
```

and example 3

```{r}
read.table("test3.txt")
```

from the slides: writing functions

## Our first function

This is an example function named `add` with input `x` and `y`

```{r}
add <- function(x, y=1) {
  # Sum the input x and y
  x + y
}
```

Let's try using it

```{r}
add(6, 4)
```
```{r}
add(x=c(1,6,2), y=4)
```
```{r}
add(c(1,6,2), 4)
```

range calculates the min and max

```{r}
x <- c(10,4,22,6)
min(x)
max(x)
range(x )
```


A 2nd example function to re-scale data to lie between 0 and 10
```{r}
rescale <- function(x) {
  rng <-range(x)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

Testing the function:

```{r}
rescale(1:10)
```

```{r}
rescale( c(1:10, 30, 100))
```
```{r}
rescale( c(1:10, NA))
```
```{r}
x <- c(1:10, NA)
rng <-  range(x)
rng
```
```{r}
x <- c(1:10, NA)
rng <-  range(x, na.rm = TRUE)
rng
```
### Fix 
```{r}
rescale2 <-  function(x) {
  rng <-  range (x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```


```{r eval=FALSE}
rescale2( c(1:10, "error example"))
```



```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
  
    rng <-range(x, na.rm=na.rm)
    print("Hello")
  
  answer <-(x - rng[1]) / (rng[2] - rng[1])
 
  print("is it me you are looking for?")
  
  if(plot) {
    print("Please don't sing again")
    plot(answer, typ="b", lwd=4)
  }
    print("I can see it in ...")
    return(answer)
}
```

```{r}
rescale3(x)
```


### This is a level 3 heading

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.




## WORKSHEET: Working with the bio3d package 

Installation is done in console; do not do it in markdown because knitting will cause it to reinstall every time.

To use the functions from any package we have installed we use `library()` function to load it.

```{r}
library(bio3d)
```

corrected the error
trims kinase down to just the A chain where drug binds (where activity is)
plots atom
try to boil it down to a function, where x is kinase ID and output is plot
```{r}
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

```{r}
s1 <- read.pdb("4AKE")
class(s1)
```

structure tells us it's a list of 8 things:
```{r}
str(s1)
```

to look at amino acid sequence
```{r}
s1$seqres
```

to change amino acid sequence to single letter abbreviation
```{r}
aa321(s1$seqres)
```

let's put it all in one plot
```{r}

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, col="blue", typ="l")
points(s3.b, col="red", typ="l", lwd=3)
```

